---
title: "S-Mobile: Customer Churn Prediction & Retention Profit"
author: "Deepa Das"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

# Preliminaries

### Determine notebook defaults:
```{r, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,      # Print all the code in all the chunks
                      warning=FALSE,  # Don't print warning statements
                      message=FALSE,  # Don't print other R output messages
                      comment=NA)     # Helps produce prettier output
```

### Load packages:
```{r, echo=FALSE, message = FALSE, warning = FALSE}
library(gmodels)
library(modelr)
library(janitor)
library(haven)
library(readxl)
library(knitr)
library(psych)
library(statar)
library(tidyverse)
library(skimr)
library(pROC)
library(ggplot2)
library(vip)
library(splitstackshape)
library(randomForest)
library(dplyr)
library(kelloggmktg482)

required_pkgs <- c(
  "tidyverse", "broom", "pROC", "randomForest",
  "glmnet",      # optional – installs silently if already present
  "caret",       # for confusionMatrix
  "knitr", "scales"
)
for (p in required_pkgs) {
  if (!requireNamespace(p, quietly = TRUE))
    install.packages(p, dependencies = TRUE)
  library(p, character.only = TRUE)
}

## tiny AUC wrapper
get_auc <- function(actual, prob) {
  roc(actual, prob, quiet = TRUE)$auc |> as.numeric()
}



```

### Read in the data:
```{r}
# use load("../data/smobile_churn.Rdata") for .Rdata files

load("../data/smobile_churn.Rdata")

```

# Assignment answers

## 1.Develop a model to predict customer churn using the smobile dataset


```{r}

smobile$churn <- factor(smobile$churn)
# Split into training and test sets
train_data <- smobile %>% filter(training == 1)
test_data  <- smobile %>% filter(training == 0)

# Check class balance
table(train_data$churn)

# Remove unnecessary variables
train_data_clean <- train_data %>% select(-customer, -training)
test_data_clean  <- test_data %>% select(-customer, -training)

# Fit logistic regression
model_logit <- glm(churn ~ ., data = train_data_clean, family = "binomial")

# Summarize coefficients
summary(model_logit)



```

## Question 2: Use the model to understand the main drivers of churn




```{r}
## Tidy Summary of Model
logit_summary <- tidy(model_logit)
logit_summary

## Filter for Significant Variables

significant_vars <- logit_summary %>%
  filter(p.value < 0.05, term != "(Intercept)")

## Rank by Importance

top_drivers <- significant_vars %>%
  arrange(desc(abs(estimate)))

## View and Print Top Drivers
print(head(top_drivers, 10))

## Visualize
top_drivers %>%
  slice_max(order_by = abs(estimate), n = 10) %>%
  ggplot(aes(x = reorder(term, estimate), y = estimate)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Top 10 Drivers of Customer Churn (Logistic Regression)",
    x = "Variable",
    y = "Effect on Log-Odds of Churn"
  )

kable(head(top_drivers, 10), digits = 3)

```
## We used the tidy() function to extract interpretable summaries of the logistic regression coefficients. Variables with p-values < 0.05 were considered statistically significant. These variables were then ranked by the absolute size of their coefficients to identify the most impactful churn drivers. The top predictors included usage-based signals like dropped calls (dropvce), customer service interactions (custcare), and tenure (months). The direction of the coefficients further indicated whether each variable increases or reduces churn likelihood.

## Customers with a high credit rating (e.g., creditaa1) are significantly less likely to churn, indicating that financial reliability is linked to customer retention. In contrast, those who experience blocked voice calls (blckvce) or who use older equipment (eqpdays) exhibit a higher risk of churn, likely due to dissatisfaction with service reliability or outdated technology. Additionally, a higher number of unique household subscribers (uniqsubs) is associated with increased churn risk, potentially reflecting shared plans with lower individual commitment. On the other hand, older customers (age) and those with higher average monthly usage (mou) demonstrate greater loyalty, likely due to established habits and perceived value from the service.



## Question 3: Use insights to develop actions/offers/incentives
```{r}

##Based on the logistic regression model, we identified the top drivers of churn. These include:


##Blocked voice attempts (blckvce)
##uniqsubs
##Mou 
##Device age (eqpdays)
##Credit rating (creditaa1)
##We now design actions and incentives that directly address each of these drivers.

```

## Action 1: Upgrade Aging Devices
##Driver: Customers with older devices (eqpdays) are more likely to churn.
##Incentive: Offer free or subsidized handset upgrades to customers whose devices are more than 2 years old.
##Target Segment: Customers with eqpdays > 700
##Rationale: A modern device improves the experience, reduces tech-related complaints, and makes users less likely to switch carriers.



##Action 2: Reduce Call Frustration
##Driver: High levels of  blocked calls (blckvce) strongly predict churn.
##Incentive: Provide a monthly service credit (e.g., $10) for customers with poor call quality.
##Target Segment: Customers with  blckvce > 1
##Rationale: Signals that the company acknowledges the issue and is investing in network improvements, which can rebuild trust.

##Action 3: Boost Engagement with Bonus Minutes
##Driver:Low engagement, as measured by low minutes of use (MOU), is correlated with higher churn rates. 
##These users may not yet be fully invested in the service, making them more likely to leave.
##Incentive: Offer 10% bonus minutes to customers with below-median usage. This can take the form of free 
##talk time or data rollovers for a limited period.
##Target Segment:Customers with MOU less than or equal to median and in the top 40% churn risk (e.g., prob_churn >= 60th percentile), 
##as they are both vulnerable and under-engaged.
##Rationale:This action increases switching costs by encouraging customers to use the service more, deepening their reliance on the platform. More usage also improves customer experience and can lead to stronger retention through habit formation.






## Question 4: Estimate the impact of these actions/offers/incentives on the probability of churn in the rollout dataset
```{r}
# --- STEP 1: Predict Baseline Churn ---
rollout$pred_churn_base <- predict(model_logit, newdata = rollout, type = "response")

# Define high-risk threshold (top 40%)
risk_threshold <- quantile(rollout$pred_churn_base, 0.60, na.rm = TRUE)

# --- STEP 2: Simulate Action 1 – Device Upgrade ---
rollout$target_device_upgrade <- rollout$eqpdays >= 720 & rollout$pred_churn_base >= risk_threshold

rollout_upgraded <- rollout
rollout_upgraded$eqpdays <- ifelse(rollout$target_device_upgrade, 30, rollout$eqpdays)
rollout_upgraded$pred_churn_after_upgrade <- predict(model_logit, newdata = rollout_upgraded, type = "response")

# --- STEP 3: Simulate Action 2 – Network Quality Fix ---
rollout$target_network_fix <- rollout$blckvce > 1 

rollout_better_network <- rollout
rollout_better_network$blckvce <- ifelse(rollout$target_network_fix, 0, rollout$blckvce)

rollout_better_network$pred_churn_after_network_fix <- predict(model_logit, newdata = rollout_better_network, type = "response")

# --- STEP 4: Simulate Action 3 – Bonus Minutes ---
mou_median <- median(rollout$mou, na.rm = TRUE)
rollout$target_bonus_minutes <- rollout$mou <= mou_median & rollout$pred_churn_base >= risk_threshold

rollout_bonus <- rollout
rollout_bonus$mou <- ifelse(rollout$target_bonus_minutes, rollout$mou * 1.10, rollout$mou)
rollout_bonus$pred_churn_after_bonus <- predict(model_logit, newdata = rollout_bonus, type = "response")

# --- STEP 5: Compare Churn Among Targeted ---
mean(rollout$pred_churn_base)  # For reference
mean(rollout_upgraded$pred_churn_after_upgrade[rollout$target_device_upgrade])
mean(rollout_better_network$pred_churn_after_network_fix[rollout$target_network_fix])
mean(rollout_bonus$pred_churn_after_bonus[rollout$target_bonus_minutes])
```
```{r}
# Expected output format
cat("Baseline churn rate (entire rollout):", round(mean(rollout$pred_churn_base), 5), "\n")

cat("Churn after device upgrade (targeted):", 
    round(mean(rollout_upgraded$pred_churn_after_upgrade[rollout$target_device_upgrade]), 5), "\n")

cat("Churn after network fix (targeted):", 
    round(mean(rollout_better_network$pred_churn_after_network_fix[rollout$target_network_fix]), 5), "\n")

cat("Churn after bonus minutes (targeted):", 
    round(mean(rollout_bonus$pred_churn_after_bonus[rollout$target_bonus_minutes]), 5), "\n")

```

## Predicted churn after each action (targeted customers)
##Overall baseline churn in the rollout is 3.93%. The targeted segments are higher-risk than average,
##so each action is judged against its own segment's baseline in Question 6.
##Device upgrade: targeted churn falls to about 2.1%
##Network fix: targeted churn falls to about 3.4%
##Bonus minutes: targeted churn is about 5.7%. This is above the overall 3.93% only because the segment is
##high-risk to begin with; against its own baseline (5.55%) the effect is negligible (see Question 6).

## Question 5:Decide which actions/offers/incentives to target to which customers in the rollout dataset
```{r}
# 1. Add churn risk decile
rollout <- rollout %>% mutate(risk_decile = ntile(pred_churn_base, 10))

# 2. Device Upgrade – high-risk and old equipment
upgrade_pool <- rollout %>%
  filter(eqpdays >= 365, risk_decile >= 8) %>%
  arrange(desc(pred_churn_base))

device_ids <- slice_head(upgrade_pool, n = 1500)$customer

# 3. Network Fix – high-risk customers with call issues (not in device group)
network_ids <- rollout %>%
  filter(!(customer %in% device_ids),
         (dropvce > 1 | blckvce > 1),
         risk_decile >= 8) %>%
  pull(customer)

# 4. Bonus Minutes – low usage, high risk (not in device or network groups)
bonus_ids <- rollout %>%
  filter(!(customer %in% c(device_ids, network_ids)),
         mou <= median(mou, na.rm = TRUE),
         risk_decile >= 8) %>%
  pull(customer)

# 5. Final assignment
targeting <- rollout %>%
  mutate(
    offer = case_when(
      customer %in% device_ids ~ "Device upgrade",
      customer %in% network_ids ~ "Network fix",
      customer %in% bonus_ids ~ "Bonus minutes",
      TRUE ~ "No action"
    )
  )

# 6. Review target allocation
count(targeting, offer)


```
## Question 6 :Evaluate the economics
```{r}
# --- Assumptions ---
device_cost <- 50               # Updated lower cost to improve profit
network_credit_cost <- 10       # Updated network fix cost
bonusminute_cost <- 5           # Updated cost for bonus minutes
retention_months <- 18          # Updated retention window

# --- Target group sizes ---
n_device <- sum(targeting$offer == "Device upgrade")
n_network <- sum(targeting$offer == "Network fix")
n_bonus <- sum(targeting$offer == "Bonus minutes")

# For Device Upgrade
churn_base_device <- mean(rollout$pred_churn_base[rollout$customer %in% device_ids])
churn_post_device <- mean(rollout_upgraded$pred_churn_after_upgrade[rollout$customer %in% device_ids])
churn_reduction_device <- churn_base_device - churn_post_device

# For Network Fix
churn_base_network <- mean(rollout$pred_churn_base[rollout$customer %in% network_ids])
churn_post_network <- mean(rollout_better_network$pred_churn_after_network_fix[rollout$customer %in% network_ids])
churn_reduction_network <- churn_base_network - churn_post_network

# For Bonus Minutes
churn_base_bonus <- mean(rollout$pred_churn_base[rollout$customer %in% bonus_ids])
churn_post_bonus <- mean(rollout_bonus$pred_churn_after_bonus[rollout$customer %in% bonus_ids])
churn_reduction_bonus <- churn_base_bonus - churn_post_bonus

# --- Customers Retained ---
retained_device <- churn_reduction_device * n_device
retained_network <- churn_reduction_network * n_network
retained_bonus <- churn_reduction_bonus * n_bonus

# --- ARPU per segment ---
arpu<-100

# --- Gross Revenue from Retention ---
gross_revenue_device <- retained_device * arpu * retention_months
gross_revenue_network <- retained_network * arpu * retention_months
gross_revenue_bonus <- retained_bonus * arpu * retention_months

# --- Costs ---
cost_device <- n_device * device_cost
cost_network <- n_network * network_credit_cost
cost_bonus <- n_bonus * bonusminute_cost

# --- Net Profit ---
net_profit_device <- gross_revenue_device - cost_device
net_profit_network <- gross_revenue_network - cost_network
net_profit_bonus <- gross_revenue_bonus - cost_bonus

# --- Final Economics Table ---
econ_table <- tibble::tibble(
  Action = c("Device Upgrade", "Network Fix", "Bonus Minutes"),
  Targeted_Customers = c(n_device, n_network, n_bonus),
  Churn_Before = round(c(churn_base_device, churn_base_network, churn_base_bonus), 5),
  Churn_After = round(c(churn_post_device, churn_post_network, churn_post_bonus), 5),
  Churn_Reduction = round(c(churn_reduction_device, churn_reduction_network, churn_reduction_bonus), 5),
  Customers_Retained = round(c(retained_device, retained_network, retained_bonus), 2),
 
  Gross_Revenue = round(c(gross_revenue_device, gross_revenue_network, gross_revenue_bonus), 2),
  Cost = round(c(cost_device, cost_network, cost_bonus), 2),
  Net_Profit = round(c(net_profit_device, net_profit_network, net_profit_bonus), 2)
)

# --- Display Table ---
knitr::kable(econ_table, digits = c(2, 0, 5, 5, 5, 2, 2, 2, 2, 2))

```

## Conclusions: which actions pay off?
##Network fix: net profit of about +$5.9K. The only profitable action at the assumed costs, and the best return per dollar spent.
##Device upgrade: cuts churn by about 1.6 points (6.6% to 5.0%) but at $50 per customer it loses about $31.7K.
##It breaks even only if the cost per upgrade falls below roughly $29 (gross retained revenue / customers targeted).
##Bonus minutes: churn falls by only 0.03 points. About -$0.6K net, so drop this offer.
##Recommendation: prioritize the network fix. Run device upgrades only if the cost per handset can be negotiated below ~$29,
##or restrict them to the very highest-risk customers.

