---
title: "Bookbinders: Customer Analysis & Statistical Testing"
author: "Deepa Das"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

# Preliminaries

## Load packages:
```{r, message = FALSE, warning = FALSE}
library(gmodels)
library(modelr)
library(janitor)
library(haven)
library(readxl)
library(knitr)
library(psych)
library(statar)
library(tidyverse)
library(vip)
library(kelloggmktg482)       # Always load last
```

## Read in the data:
```{r}
# use load("../data/filename.Rdata") for .Rdata files
load("../data/bbb.Rdata")
```

# Assignment questions and answers

## 1. Previously, you reported the correlation between customers' total spending on non-book products and on books (see the `nonbook` and `book` variables). Now, report a 95% confidence interval for this correlation.
```{r}
cor.test(bbb$nonbook,bbb$book)
# The correlation co-efficient is 0.1574
# 95 % confidence interval is 0.1488 to 0.1659

```

## 2. Report the output of a linear regression predicting customers' total spending on non-book products from their spending on book products as well as 95% confidence intervals for the intercept and slope.
```{r}
lin_reg<-lm(formula=nonbook~book,data=bbb)
# View regression output
summary(lin_reg)

# Get 95% confidence intervals for intercept and slope
confint(lin_reg)
#The estimate for intercept and slope are 144.47388 and 0.36331
#the 95% confidence interval for intercept is 143.2670547 145.6807102
#the 95% confidence interval for slope is 0.3433358   0.3832878

```

## 3. Previously, you reported the proportion of customers by gender who bought "The Art History of Florence"  (see the `buyer` variable). Now, report a 95% confidence interval for the difference in the proportion of buyers between genders.
```{r}
library(dplyr)

counts <- table(bbb$gender, bbb$buyer)
buyers <- counts[,"1"]
totals <- rowSums(counts)
prop <- prop.test(buyers, totals, correct=FALSE)
prop
prop$conf.int
#95% confidence interval for the difference in the proportion of buyers between genders 
#is -0.06177431 to -0.05023053
```

## 4. Report the results of a chi-square test of the association between the state in which the customer lives and whether or the customer bought "The Art History of Florence".
```{r}
chisq.test(bbb$state,bbb$buyer,correct=FALSE)
#The chi-squared test result is 23.549 with a p value of 0.0519
#Since p value is greater than 0.05, but very close to 0.05 , it suggests that the reltionship 
#between state and whether they bought "The art history of florence" is less statistically significant 
#additional analysis may be needed to assess the significance
```


## 5. Report the result of a t-test comparing the average number of purchases by gender as well as a 95% confidence interval for the difference in the average number of purchases between genders.
```{r}

purch_gender<-t.test(purch~gender,data=bbb,var.equal=TRUE)
print(purch_gender)
#The mean for females is 3.362200 and the mean for males is 4.943287 
# The 95% confidence interval for diferrence in the average number of purchases 
#is  -1.644191 -1.517981

```


## 6. Repeat the same analysis but using a linear regression rather than a t-test.
```{r}
purchase_bygender<-lm(formula(purch~gender),data=bbb)
summary(purchase_bygender)
confint(purchase_bygender)
#the null hypothesis is rejected, p value is 2e-16
#there is statistically significant relation between average number ofpurchase
#and gender

```

## 7. In  class, we examined a logistic regression predicting the `buyer` variable. In this assigment, we want to predict whether the customer is female versus male. First, we must create `female` binary variable using the following:
```{r}
bbb <- bbb %>% mutate(female = factor(1 * (gender=="F")))

```
## Before proceeding to fit a logistic regression, first provide graphical evidence that the variables `do_it`, `total`, and `last` are or are not predictive of whether the customer is female versus male.
```{r}
library(tidyverse)

# Boxplot for do_it
ggplot(bbb, aes(x = gender, y = do_it)) +
  geom_boxplot() +
  labs(title = "Do-it Spending by Gender", y = "Do-it", x = "Gender")

# Boxplot for total spending
ggplot(bbb, aes(x = gender, y = total)) +
  geom_boxplot() +
  labs(title = "Total Spending by Gender", y = "Total", x = "Gender")

# Boxplot for months since last purchase
ggplot(bbb, aes(x = gender, y = last)) +
  geom_boxplot() +
  labs(title = "Months Since Last Purchase by Gender", y = "Last", x = "Gender")

# do_it is a strong predictor: men clearly purchase more DIY books than women.

#total variable:  Male customers have slightly higher median total spending than 
#females.The upper quartile for men is slightly more spread out than for women.
#Still, there's a lot of overlap.So this variable is a weak predictor

#The median is about the same for both males and females (~11–12 months).
#The spread is similar for both genders.No meaningful difference in recency of purchase 
#between genders.So, 'last' is not a strong predictor of gender.

```

## 8. Report the output of a logistic regression predicting `female` using the variables `last`, `total`, `child`, `youth`, `cook`, `do_it`, `reference`, `art`, and `geog`  as well as 95% confidence intervals for the intercept and coefficients.
```{r}
log_model <- glm(female ~ last + total + child + youth + cook + do_it + reference + art + geog,
                 data = bbb,
                 family = "binomial")
summary(log_model)
confint(log_model)

#last	- co-efficient is -4.903e-04 and 95% CI is	(-0.0029, 0.0019)	
# last is not significant predictor of gender

#total- co-efficient is	9.244e-05 and 95% CI is	(-0.00013, 0.00032)	0.420	
#total is not significant  predictor of gender

#child- co-efficient is	-3.582e-02  and 95% CI is	(-0.055, -0.017)
#significant predictor of female

#youth- co-efficient is	-1.137e-02 and 95% CI is	(-0.041, 0.019)	0.456	
#youth is not a significant predictor of female

#cook	- co-efficient is -2.723e-02 and 95% CI is	(-0.046, -0.009)
#cook is a significant predictor 

#do_it- co-efficient is	-7.634e-01 and 95% CI is	(-0.792, -0.735)	
#significant predictor

#reference- co-efficient is	-9.022e-02 and 95% CI is	(-0.123, -0.057)
#Significant predictor

#art- co-efficient is	-8.414e-02 and 95% CI is	(-0.114, -0.054)
#Significant predictor

#geog- co-efficient is -7.399e-02	 and 95% CI is	(-0.099, -0.049)	
#Significant predictor

#Summary:The variables, do_it, child, cook, reference, art, and geog, are 
#significant predictors of the 'gender' variable, since their p-values are below
#0.05.
```

## 9. Report the variable importance for all variables included in the logistic regression.
```{r}
varimpplot(log_model,target="female",data =bbb)
#consistent with Question 8 interpretation.The variables, do_it, child, cook, 
#reference, art, and geog have impact on the target variable.
#x-axis is in RMSE units
#do_it has approx.11% impact when omitted in the target variable prediction
#geog has  approx.2% impact when omitted in the target variable prediction
#art has  approx.1.7% impact when omitted in the target variable prediction
#reference has  approx.1.6% impact when omitted in the target variable prediction
#child has  approx.0.75% impact when omitted in the target variable prediction
#cook has  approx.0.6% impact when omitted in the target variable prediction
#youth has  approx.0% impact when omitted in the target variable prediction
#last has  approx.0% impact when omitted in the target variable prediction
#total has  approx.0% impact when omitted in the target variable prediction
```

## 10. Report partial dependence plots for all variables included in the logistic regression. Comment on whether the partial dependence plots for `do_it`, `total`, and `last` are consistent or inconsistent with those you found in Question 7 and explain any consistency or inconsistency.
```{r}
female.overall<-mean(bbb$female)
pardepplot(log_model,pred.var="do_it",data=bbb,hline=female.overall)
pardepplot(log_model,pred.var="geog",data=bbb,hline=female.overall)
pardepplot(log_model,pred.var="art",data=bbb,hline=female.overall)
pardepplot(log_model,pred.var="reference",data=bbb,hline=female.overall)
pardepplot(log_model,pred.var="child",data=bbb,hline=female.overall)
pardepplot(log_model,pred.var="cook",data=bbb,hline=female.overall)
pardepplot(log_model,pred.var="youth",data=bbb,hline=female.overall)
pardepplot(log_model,pred.var="last",data=bbb,hline=female.overall)
pardepplot(log_model,pred.var="total",data=bbb,hline=female.overall)

# #The partial dependence plot for do_it is consistent with the relationship 
#deduced from the boxplot in question 7. The mean spending on do_it (DIY books) 
#is higher for gender 'M'. Thus, in the partial dependence plot as do_it 
#increases the predicted variable 'female' decreases.

#As total spending increases, the predicted probability of being female also 
#increases.However, the slope is very shallow.At $0 spending, phat is approx 0.662
#At $500 spending, phat is approx 0.671.This is a total increase of less than 1% point.
#This finding is consistent with the logistic regression results and boxplot in 
#Q7, which showed no meaningful difference in total spending by gender. While the 
#slope is technically positive, the strength of the relationship is so weak that 
#total remains an unimportant predictor between M and F customers.

#the partial dependence plot for 'last' is also consistent with
#question 7, since as per the output in question the mean number of months since
#last purchase was approximately the same level for both genders and here the
#probability of being female (gender F) is decreasing by only a very small 
#amount with increase in the number of months since last purchase. The partial 
#dependence plot for last shows a very weak negative trend, suggesting that 
#customers who haven’t purchased recently are slightly less likely to be female. 
#However, the effect size is extremely small, with predicted probabilities 
#changing by less than 0.004 across the full range of values. This confirms that
#last is not a meaningful predictor of gender, consistent with the boxplot in 
#Question 7 and its low importance in the regression model.Since the
#decrease observed in the plot is only 0.01.

```

## 11. Add the predicted probabilities from the logistic regression to the bbb dataframe. Then report histograms of the predicted probabilities separately for each value of `do_it`. Comment on whether the results are consistent with the results of the partial dependence plots from Question 10. 
Hint: To accomplish this, recall facet_wrap() from the R tutorial "the R tutorial." You may want to use "scales='free_y'" within facet_wrap() * because relatively few people bought a larger number of do-it-yourself books.
[* or aes(y=..density..) within geom_histogram()]
```{r}
# Add predicted probabilities (of being female) to the dataset
bbb$predicted_prob <- predict(log_model, type = "response")

library(ggplot2)

ggplot(bbb, aes(x = predicted_prob)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  facet_wrap(~ do_it, scales = "free_y") +
  labs(
    title = "Predicted Probability of Being Female by Do-It Book Count",
    x = "Predicted Probability (Female)",
    y = "Count"
  ) 


#Top-left panel:This group is huge and most customers bought no DIY books.
#Predicted probabilities are clustered around 0.6–0.8, meaning,
#these customers are much more likely to be female.
#Distribution shifts gradually leftward as do_it increases.
#For do_it = 3, most probabilities are around 0.2–0.3.
#As customers buy more DIY books,model becomes less confident of being female
#this is consistent with partial dependence plot in Q10



```

## 12. Report histograms of `total` separately for each value of `do_it`. Does this plot help explain consistency or inconsistency discussed in Question 10?
```{r}
library(ggplot2)

ggplot(bbb, aes(x = total)) +
  geom_histogram(bins = 30, fill = "darkgreen", color = "white") +
  facet_wrap(~ do_it, scales = "free_y") +
  labs(
    title = "Distribution of Total Spending by Do-It Book Count",
    x = "Total Spending",
    y = "Count"
  ) 
# The histogram shows that total spending overlaps substantially across all levels of 
#`do_it`.
# While higher `do_it` values are associated with somewhat higher spending ranges, 
# customers with low `do_it` also span a wide range of total spending.
# Additionally, the number of customers decreases dramatically as `do_it` increases.
# This overlap explains why `total` is a poor predictor of gender in the logistic regression, 
# even though `do_it` is a strong one.
# This pattern is consistent with the partial dependence plots in Q10: 
# the PDP for `do_it` showed a steep drop in probability of being female, 
# while the PDP for `total` showed only a shallow increase. 
# Therefore, the results here support and help explain the findings in Q10.

```